---
title: "simulate-species"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulate-species}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

An example of how to the Human-Wildlife Conflict Index suite works - simulate some data - collate sets to represent example species times - calculate and explore index suite for species and landscape levels.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hwci.tools)
library(tidyverse)
library(patchwork)
```

## Simulate some data

Timeline
```{r}
nyears <- 100

time_series <- 1:nyears
```

Landscape population and number of households
```{r}
#landscape population
k <- 500

# average household size
l <- 5

# number of households in landscape
j <- k/l
```

Human-victim incidents
```{r}
v_mu <- 20

v_none   <- rep(0, times = nyears)
v_low    <- rnbinom(n = nyears, mu = v_mu,                   size = v_mu/2)
v_rising <- rnbinom(n = nyears, mu = v_mu + 0.2*time_series, size = v_mu/2)
v_high   <- rnbinom(n = nyears, mu = v_mu * 3,               size = v_mu/2)
```

Human mortalities
```{r}
m_mu <- 5

m_none <- rep(0, times = nyears)
m_low    <- rnbinom(n = nyears, mu = m_mu,                    size = m_mu/2) %>%
  if_else(v_low < ., v_low, .)
m_rising <- rnbinom(n = nyears, mu = m_mu + 0.05*time_series, size = m_mu/2) %>%
  if_else(v_rising < ., v_rising, .)
m_high   <- rnbinom(n = nyears, mu = m_mu * 3,                size = m_mu/2) %>%
  if_else(v_high < ., v_high, .)
```

Economic incidents
```{r}
b_mu <- 15

b_none <- rep(0, times = nyears)
b_low    <- rnbinom(n = nyears, mu = b_mu,                    size = b_mu/2)
b_rising <- rnbinom(n = nyears, mu = b_mu + 0.15*time_series, size = b_mu/2)
b_high   <- rnbinom(n = nyears, mu = b_mu * 4,                size = b_mu/2)
```

Value of economic incidents
```{r}
l_mu <- 100

l_none <- vector("list", length = nyears)
l_none[1:nyears] <- 0

l_flat_b_low <- lapply(b_low, FUN = function(x){
  rnbinom(n = x, mu = l_mu, size = l_mu/20)
})

l_rising_b_low <- lapply(b_low, FUN = function(x){
  rnbinom(n = x, mu = l_mu + time_series, size = l_mu/20)
})

l_rising_b_low <- mapply(
  FUN = function(x, y){
    rnbinom(n = x, mu = l_mu + y, size = l_mu/20)
  },
  x = b_low,
  y = time_series
)

l_flat_b_rising <- lapply(b_rising, FUN = function(x){
  rnbinom(n = x, mu = l_mu, size = l_mu/20)
})

l_rising_b_rising <- mapply(
  FUN = function(x, y){
    rnbinom(n = x, mu = l_mu + y, size = l_mu/20)
  },
  x = b_rising,
  y = time_series
)
```

Household wealth (mean)
```{r}
w_bar <- 1000
```

Wildlife-victim incidents
```{r}
r_mu <- 5

r_none <- rep(0, times = nyears)
r_low    <- rnbinom(n = nyears, mu = r_mu,                    size = r_mu/2) %>%
  if_else(v_low < ., v_low, .)
r_rising <- rnbinom(n = nyears, mu = r_mu + 0.05*time_series, size = r_mu/2) %>%
  if_else(v_rising < ., v_rising, .)
r_high   <- rnbinom(n = nyears, mu = r_mu * 3,                size = r_mu/2) %>%
  if_else(v_high < ., v_high, .)

```

[IUCN-Red List status](https://www.iucnredlist.org/resources/categories-and-criteria)
```{r}
s_cr <- "CR"
s_en <- "EN"
s_vu <- "VU"
s_nt <- "NT"
s_lc <- "LC"
```


## Create data sets for example species 

Species 1: large carnivore type - human victims, no economic damage, reprisals
```{r}
sp1_data <- hwci_data(
  t = time_series,
  j = j,
  v = v_rising,
  m = m_rising,
  b = b_none,
  l = l_none,
  w = w_bar,
  r = r_rising,
  s = "CR"
)
sp1_data
```

Species 2: crop pest type - no human incidents, only economic incidents
```{r}
sp2_data <- hwci_data(
  t = time_series,
  j = j,
  v = v_none,
  m = m_none,
  b = b_rising,
  l = l_flat_b_rising,
  w = w_bar,
  r = r_none,
  s = "LC"
)
```

Species 3: elephant type - human and economic incidents, incidents getting worse, few reprisals
```{r}
sp3_data <- hwci_data(
  t = time_series,
  j = j,
  v = v_rising,
  m = m_rising,
  b = b_rising,
  l = l_rising_b_rising,
  w = w_bar,
  r = r_low,
  s = "EN"
)
```


## Calculate species indices
```{r}
sp1_indices <- hwci(data = sp1_data)
```


```{r}
sp2_indices <- hwci(data = sp2_data)
```


```{r}
sp3_indices <- hwci(data = sp3_data)
sp3_indices
```

## Calculate landscape metrics
```{r}
allsp_lm <- hwci_lm(sp1_indices, sp2_indices, sp3_indices)
allsp_lm
```

